\section{Maths Jokes}

\begin{itemize}
 \item
  Q: What do you call something that's purple and equipped with a commutative
  operation? \\
  A: An Abelian grape!
 \item
  Q: What sound does a drowning number theoretician make? \\
  A: \(\log \log \log \log \log \log \log\)
 \item
  A countably infinite number of mathematicians walk into a bar.
  {The first asks for one pint.
  \footnotesize The second asks for half a pint.
  \scriptsize The third asks for a quarter of a pint.
  \tiny The fourth asks for an eighth of a pint.}
  The bartender pours two pints and proclaims, ``Know your limits!''
 \item
  Q: What did \(i \pi\) say to the exponential function? \\
  A: You make me feel mighty real.
 \item
  Q: What is the result of evaluating the indefinite integral
  \begin{equation*}
   \integ{\frac 1{\mathrm{cabin}}}{\,\mathrm{cabin}} \ \text?
  \end{equation*}
  A (hopefully): *exasperated sigh* god you're an idiot, it's a
  \(\log \mathrm{cabin}\).

  Jubilant rebuttal: Actually, it's a houseboat! You forgot to add the C!
 \item
  An engineer, a physicist, and a mathematician are shown a pasture with a herd
  of sheep, and told to put them inside the smallest possible amount of fence.
  \\
  The engineer is first. He herds the sheep into a circle and then puts the
  fence around them, declaring, ``A circle will use the least fence for a given
  area, so this is the best solution.'' \\
  The physicist is next. He creates a circular fence of infinite radius around
  the sheep, and then draws the fence tight around the herd, declaring, ``This
  will give the smallest circular fence around the herd.'' \\
  The mathematician is last. After giving the problem a little thought, he puts
  a small fence around himself and then declares, ``I define myself to be on the
  outside.''
 \item
  An engineer, a physicist, and a mathematician were on a train heading north,
  and had just crossed the border into Scotland. \\
  The engineer looked out of the window and said ``Look! Scottish sheep are
  black!'' \\
  The physicist said, ``No, no. Some Scottish sheep are black.'' \\
  The mathematician looked irritated. ``There is at least one field, containing
  at least one sheep, of which at least one side is black.'' \\
 \item
  An engineer and a mathematician were shown into a kitchen, given an empty pan,
  and told to boil a pint of water. They both filled the pan with water, put it
  on the stove, and boiled it. \\
  The next day they were shown into the kitchen again, given a pan full of
  water, and told to boil a pint of water. \\
  The engineer took the pan, put it on the stove, and boiled it. \\
  The mathematician took the pan and emptied it, thereby reducing it to a
  previously solved problem.
 \item
  A mathematician, physicist, and engineer are taking a math test. One question
  asks ``Are all odd numbers prime?'' \\
  The mathematician thinks, ``3 is prime, 5 is prime, 7 is prime, 9 is not prime
  -- nope, not all odd numbers are prime.'' \\
  The physicist thinks, `` 3 is prime, 5 is prime, 7 is prime, 9 is not prime --
  that could be experimental error -- 11 is prime, 13 is prime, yes, they're all
  prime.'' \\
  The engineer thinks, `` 3 is prime, 5 is prime, 7 is prime, 9 is prime, 11 is prime, ...''
 \item
  A physicist, a biologist and a mathematician sit in a sidewalk cafe, looking
  at the building across the road. Two people go into the building, then three
  people come out. \\
  Physicist: ``This must be a measuring error!'' \\
  Biologist: ``This is proof of procreation!'' \\
  Mathematician: ``If one more person goes into the building, it will be
  empty!''
\end{itemize}
