% This is a LaTeX document, compiled with pdflatex.
%
% Sometimes the comments in this document contain hints or partial answers for
% questions. If you wish not to be spoiled, do not read the source in its
% entirety.

\documentclass[fleqn,a4paper,11pt]{article}
\date{}
\author{}
\title{Additional maths questions (draft ed.)}

\usepackage{mymaths}
\usepackage{mysty}

\begin{document}
\maketitle
\tableofcontents

\section{Introduction}

This is a collection of bits of mathematics relating to the Further Maths
A-level and beyond. My alternative click-bait idea for a title was ``What they
don't teach you in Further Maths A-level''. This document is often concerned
much more with proof than you would normally see at A-level, but I think this is
a very important part of mathematics (and to me, it makes maths much easier to
really agree with).

In some cases, this document may well take things way too fast. You can always
wait until you've met more things at A-level before tackling eg some of the
tricker integration questions.

Also there are certainly questions here that you won't be able to answer, at
least not straight away (I hope). I'm not very good at gauging difficulty of
questions that I've written, so I won't always know which ones these are. If a
question seems just completely impossible, you can often just leave it.
Sometimes this will prohibit you from doing some other later questions, and if
so you will probably notice this and you can leave those too. It is worthwhile
to just let such questions simmer and see if you can come to an answer by
yourself eventually.

Unless a question is modulated by a phrase like ``briefly explain'', the
expectation is to provide sound mathematical arguments as answers. Of course
it's up to you how you write down your answers, but you should at least be
convinced that you could make your answers rigorous if you wanted to. Initial
scribblings never have to fulfil any arbitrary standard, but it can be useful to
then ``neaten up'' your answer once you have one.

\input{tex/notation}
\input{tex/logic}
\input{tex/functions_sets}
\input{tex/number_theory}
\input{tex/trigonometry}
\input{tex/calculus}
\input{tex/complex}
\input{tex/matrices}
\end{document}
