\section{Complex Numbers}

\begin{enumerate}
 \item
  Show that addition and multiplication in the complex numbers are commutative
  and associative.
 \item
  Show that multiplication distributes over addition in the complex numbers.

  (This means that \(a(b + c) = ab + ac \Forall a, b, c \in \Complex\))
 \item
  Show that if \(x\) is real, then \(x(z_1, z_2) \equiv (x z_1, x z_2)\).
 \item
  Show that if \(x, y \in \Reals\), then \((x, y) \equiv x + i y\).

  This gives rise to the conventional notation \(x + iy\) for an arbitrary
  complex number. That is, if you are given \(z \in \Complex\), it fine to say
  ``let \(x, y \in \Reals\) be such that \(z = x + iy\).''
 \item
  Show that \(i^2 = -1\).
 \item
  Show that if \(z, w \in \Complex\), then \(\abs{zw} = \abs z \abs w\).
  % |(x + iy)(t + is)| = |xt - ys + i(yt + xs)|
  %                    = √((xt - ys)^2 + (yt + xs)^2)
  %                    = √(x²t² + y²s² - 2xtys + y²t² + x²s² + 2ytxs)
  %                    = √(x²t² + y²s² + y²t² + x²s²)
  %                    = √((x² + y²)(t² + s²))
 \item
  If \(x, y \in \Reals\) and \(z = x + iy\), find a \(w \in \Complex\) such that
  \(zw = \abs z^2\). \(w\) is usually written \(\conj z\).

  If \(z \ne 0\), find a complex number \(z^{-1} \in \Complex\) such that
  \(zz^{-1} = 1\). Write it in terms of \(z\) and \(\conj z\).
 \item \label{q_compl_args}
  Given two complex numbers
  \begin{align*}
   z &= r_1(\cos \theta_1 + i \sin \theta_1) \\
   w &= r_2(\cos \theta_2 + i \sin \theta_2)
  \end{align*}
  where \(r_1, r_2, \theta_1, \theta_2 \in \Reals\),
  calculate the values of \(\abs z\) and \(\abs w\).

  Draw a diagram in the complex plane\footnote{
   The complex plane is a two-dimensional plane with ``real part'' running along
   the horizontal axis, and ``imaginary part'' running along the vertical axis,
   used to represent complex numbers.
  } to interpret the values of \(r_1\), \(r_2\), \(\theta_1\), and \(\theta_2\).

  Calculate \(zw\) and write it in the form
  \(R_1 \cos \phi_1 + R_2 \sin \phi_2\). Hence calculate \(\abs{zw}\) and draw
  \(zw\) on your diagram.

  Could \(z\) be any complex number? Write down a pair of values
  \((r_1, \theta_1)\) so that \(z = -1 - \sqrt 3i\).

  Write down another such pair of values for which \(r_1\) is the same but
  \(\theta_1\) is different. Also write down another pair for which
  \(\theta_1\) is the same but \(r_1\) is different.
 \item
  Carefully write down the first eight terms of the Maclaurin series of
  \(e^{i\theta}\) (by expanding \(e^x\) where \(x = i\theta\)). By comparing
  with the Maclaurin series of \(\sin \theta\) and \(\cos \theta\), write down a
  sensible formula for \(e^{i\theta}\), of the form
  \(\alpha \cos \theta + \beta \sin \theta\) where
  \(\alpha, \beta \in \Complex\).
 \item
  You may now assume that this formula is correct\footnote{
   Really, this is just how we choose to define \(e^z\) for \(z \in \Complex\),
   so its correctness is literally trivial.
  }.

  Use question \ref{q_compl_args} to show that for \(\theta, \phi \in \Reals\),
  \(e^{i\theta} e^{i\phi} \equiv e^{i(\theta + \phi)}\).
 \item
  For \(z = x + iy \in \Complex\), we now further define
  \(e^z \defeq e^x e^{iy}\).

  Hence prove by induction that for \(z \in \Complex\), \(n \in \Naturals\),
  \((e^z)^n \equiv e^{nz}\).

  Write down the values of \(e^{i\pi}\), \(e^{i\pi / 2}\),
  \(e^{i\pi / 4}\), \(e^{i\pi / 3}\), and \(e^{2 i \pi}\).
 \item
  Use the formula to express
  \((\cos \theta + i \sin \theta)^n\) in the form \(\cos \phi + i \sin \phi\).

  This is called ``de Moivre's Theorem''.
 \item
  By considering the real part of \(e^{5ix}\), show that
  \begin{equation*}
   \cos 5x \equiv 16\cos^5 x - 20\cos^3 x + 5\cos x
  \end{equation*}
  You may wish to write \(c = \cos x\) and \(s = \sin x\) for convenience.
  % cos 5x = Re((c + is)^5)
  %        = c^5 - c^3 s^2 + c s^4
  %        = c^5 - 10 c^3 (1 - c^2) + 5 c (1 - c^2)^2
  %        = c^5 - 10c^3 + 10c^5 + 5c - 10c^3 + 5c^5
  %        = 16c^5 - 20c^3 + 5c
 \item
  Show that if \(n \in \Naturals\),
  \begin{align*}
   \cos 2nx &\equiv
    \sum_{r = 0}^{n - 1}
     (-1)^r \binom{2n}{2r} \cos^{2(n - r)} x \sin^{2r} x \\
   \sin 2nx &\equiv
    \sum_{r = 0}^{n - 1}
     (-1)^r \binom{2n}{2r + 1} \cos^{2(n - r) - 1} x \sin^{2r + 1} x
  \end{align*}
  and deduce that
  \begin{equation*}
   \tan 2nx \equiv
   \frac{\sum_{r = 0}^{n - 1} (-1)^r \binom{2n}{2r + 1} \tan^{2r + 1} x}
        {\sum_{r = 0}^{n - 1} (-1)^r \binom{2n}{2r} \tan^{2r} x}
  \end{equation*}
  % cos 2nx + i·sin 2nx = (c + is)^2n
  %                    = Σ (2n)Cr c^(2n - r) (is)^r
  %                    = Σ (2n)Cr c^(2n - r) i^r s^r
  %                    = Σ (2n)C(2r) c^(2n - 2r) i^2r s^2r +
  %                      Σ (2n)C(2r + 1) c^(2n - 2r - 1) i^(2r + 1) s^(2r + 1)
  %                    = Σ (2n)C(2r) c^(2n - 2r) (-1)^2r s^2r +
  %                     iΣ (2n)C(2r + 1) c^(2n - 2r - 1) (-1)^r s^(2r + 1)
 \item
  Write down \(\alpha_1, \alpha_2, \beta_1, \beta_2 \in \Complex\) such that
  \begin{align*}
   e^{ix} &= \alpha_1 \cos x + \beta_1 \sin x \\
   e^{-ix} &= \alpha_2 \cos x + \beta_2 \sin x
  \end{align*}
  Solve this system of equations to find nice formulae for
  \(\sin x\) and \(\cos x\) in terms of \(e^{\pm ix}\).
 \item
  Hence find a solution \(z \in \Complex\) to \(\sin z = 2\).

  Is this the only solution?
  %      (e^ix - e^-ix)/2i = 2
  % ⇔          e^2ix - 1 = 4ie^ix
  % ⇔ e^2ix - 4ie^ix - 1 = 0
  % ⇔               e^ix = (4i ± √(-12)) / 2
  %                      = i(2 ± √3)
  %                      = e^(iπ/2 + log(2 ± √3))
  % ⇔                 iz = iπ(1/2 + 2n) + log(2 ± √3)
  % ⇔                  z = π(1/2 + 2n) - i·log(2 ± √3)
 \item
  We write \(\Integers[i]\) to denote the Gaussian integers, which is the set of
  complex numbers with integer real and imaginary parts - ie
  \(\set{a + bi : a, b \in \Integers}\).

  Let \(n \in \Integers[i]\). Show that \(\abs n^2 \in \Integers\).

  Explain why \(n^2\) corresponds to a Pythagorean triple.
 \item \label{q_comp_zw_exp}
  Spot the mistake in the reasoning:
  \begin{tcolorbox}
   \begin{align*}
    (-1)^{1/2} (-1)^{1/2}
     &= (e^{i\pi})^{1/2} (e^{i\pi})^{1/2} \\
     &= e^{i\pi/2} e^{i\pi/2} \\
     &= -1
   \end{align*}
   but also,
   \begin{align*}
    (-1)^{1/2} (-1)^{1/2}
     &= ((-1) \cdot (-1))^{1/2} \\
     &= 1^{1/2} \\
     &= (e^{0 \cdot i})^{1/2} \\
     &= e^0 \\
     &= 1
   \end{align*}
   so \(-1 = 1\).
  \end{tcolorbox}
 \item
  Find a proper subset \(G \subset \Complex\) satisfying the following
  properties:
  \begin{itemize}
   \item
    There is a \(z \in G\) with \(\Re(z) \ne 0\).
   \item
    There is a \(z \in G\) with \(\Im(z) \ne 0\).
   \item
    For all \(z, w \in G\), we have \(z  + (-1 \cdot w) \in G\).
  \end{itemize}
  Is there such a \(G\) that is finite?
\end{enumerate}

\subsection{An aside on complex exponentiation}

I will try to convince you here that our definition for \(e^{ix}\) really is the
right one. If we are to define \(e^{ix}\) in \(\Complex\), then we would like
two things: \(e^{i \cdot 0} = 1\), and \(\dv{x} (e^{ix}) = ie^{ix}\). The latter
is \emph{crucial} as the exponential function is so interlinked with calculus.
An exponential that we could not differentiate would hardly be an exponential
function at all.

But since multiplication by \(i\), aka
\(\cos \frac \pi 2 + i \sin \frac \pi 2\) is just anti-clockwise rotation by
\(\degrees{90}\) in the complex plane (cf question \ref{q_compl_args}), this
means that if we consider the point tracing out \(e^{it}\) as \(t\) varies its
``velocity'' is always perpendicular to its ``position'' with respect to the
origin. Furthermore, we can differentiate again to get
\(\dv[2]{x} (e^{ix}) = -e^{ix}\), so its ``acceleration'' is always towards the
origin. Now, if you have done some physics, this should start to sound familiar.
This is just a simple stable circular orbit!

If you trust differential equations, suppose that
\(e^{it} \equiv x(t) + i y(t)\). Then \(x'' = -x\) and \(y'' = -y\), and we have
the initial conditions \(x(0) = 1\), \(x'(0) = 0\), and \(y(0) = 0\),
\(y'(0) = 1\). These are sufficient to show that
\(x(t) = \cos t\) and \(y(t) = \sin t\).

It is then \emph{devilishly} difficult to define \(z^w\) satisfactorily for
general \(z, w \in \Complex\). It is tempting to just say something like ``write
z in the form \(re^{i\theta}\) where \(r \in \Reals^+\) and
\(\theta \in \intco{0, 2\pi}\), and define
\(z^w \defeq e^{(\ln r + i\theta)w}\)''.
However, then useful identities like \((zw)^c = z^c w^c\) start to fail (cf
question \ref{q_comp_zw_exp}).

(Interesting side note: almost whatever definition you \emph{do} use, \(i^i\)
will be real).

A way to resolve this is to just call \(z^w\) a ``multivalued'' function
(violating everything we know and hold dear about functions), that takes on
the values corresponding to each \(\theta \in \Reals\), rather than just the
\(\theta \in \intco{0, 2\pi}\).
